
\documentclass[10pt,twocolumn,letterpaper]{article}
% Include other packages here, before hyperref.
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{verbatim}

% \usepackage{xcolor}         % colors
\usepackage{subfigure}

\usepackage{wrapfig}
\usepackage{times}
\usepackage{epsfig}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
%%%%%%%%% PAPER TYPE  - PLEASE UPDATE FOR FINAL VERSION
% \usepackage[review]{cvpr}      % To produce the REVIEW version
%\usepackage{cvpr}              % To produce the CAMERA-READY version
\usepackage[pagenumbers]{cvpr} % To force page numbers, e.g. for an arXiv version


\usepackage{siunitx}
\usepackage{cite}
\usepackage{ctable}
\usepackage{hhline}
\usepackage{booktabs}

\usepackage{csquotes}
\usepackage{pifont}
\usepackage{mathtools}
\usepackage{animate}
\newlength{\itemwidth}
\usepackage{anyfontsize}
\usepackage[super]{nth}

\usepackage{ctable}
\usepackage{soul}
\usepackage{multicol}
\usepackage{makecell}
\usepackage[normalem]{ulem}
\usepackage{multirow}
\usepackage{float}
\usepackage{stfloats}

\usepackage{stmaryrd}
\usepackage{mathstyle}

\definecolor{Gray}{gray}{0.9}

\newcommand{\misscite}{\textcolor{red}{[C]~}}
\newcommand{\missref}{\textcolor{red}{[R]~}}
\newcommand{\missvalue}{\textcolor{red}{[V]~}}

\providecommand{\ie}{\textit{i.e.}\@\xspace}
\providecommand{\eg}{\textit{e.g.}\@\xspace}
\providecommand{\etal}{\textit{et al}\@\xspace}

\usepackage{xcolor}         % colors
\usepackage{color, colortbl}
\definecolor{red}{RGB}{178,34,34}
\definecolor{blue}{RGB}{0,0,255}
\newcommand{\TODO}[1]{\textcolor{red}{(TODO: #1)}}
\usepackage{xspace}
\usepackage{mmstyle}
\usepackage{multirow}
 \usepackage{threeparttable} 
\usepackage{subfigure} % 子图包
% It is strongly recommended to use hyperref, especially for the review version.
% hyperref with option pagebackref eases the reviewers' job.
% Please disable hyperref *only* if you encounter grave issues, e.g. with the
% file validation for the camera-ready version.
%
% If you comment hyperref and then uncomment it, you should delete
% ReviewTempalte.aux before re-running LaTeX.
% (Or just hit 'q' on the first LaTeX run, let it finish, and you
%  should be clear).
\usepackage[pagebackref,breaklinks,colorlinks]{hyperref}

% Support for easy cross-referencing
\usepackage[capitalize]{cleveref}
\crefname{section}{Sec.}{Secs.}
\Crefname{section}{Section}{Sections}
\Crefname{table}{Table}{Tables}
\crefname{table}{Tab.}{Tabs.}



\newcommand{\modelname}{\emph{LTSF-Linear}\xspace}
\begin{document}


\title{Are Transformers Effective for Time Series Forecasting?}

\newcommand*{\affaddr}[1]{#1} % No op here. Customize it for different styles.
\newcommand*{\affmark}[1][*]{\textsuperscript{#1}}
\newcommand*{\email}[1]{\texttt{#1}}

\makeatletter
\newcommand{\printfnsymbol}[1]{%
  \textsuperscript{\@fnsymbol{#1}}%
}
\makeatother

\author{%
Ailing Zeng\affmark[1]\thanks{Equal contribution} , Muxi Chen\affmark[1]\printfnsymbol{1}, Lei Zhang\affmark[2], Qiang Xu\affmark[1]\\
\affaddr{\affmark[1]The Chinese University of Hong Kong}\\
\affaddr{\affmark[2]International Digital Economy Academy (IDEA)}\\
\email{\tt\small\{alzeng, mxchen21, qxu\}@cse.cuhk.edu.hk}\\
\email{\tt\small\{leizhang\}@idea.edu.cn}
}



% REMOVE THIS: bibentry
% This is only needed to show inline citations in the guidelines document. You should not need it and can safely delete it.
%\usepackage{bibentry}
% END REMOVE bibentry


\maketitle

\begin{abstract}
Recently, there has been a surge of Transformer-based solutions for the long-term time series forecasting (LTSF) task. Despite the growing performance over the past few years, \emph{we question the validity of this line of research in this work}.
Specifically, Transformers is arguably the most successful solution to extract the semantic correlations among the elements in a long sequence. However, in time series modeling, we are to extract the temporal relations in \emph{an ordered set of continuous points}. While employing positional encoding and using tokens to embed sub-series in Transformers facilitate preserving some ordering information, the nature of the \emph{permutation-invariant} self-attention mechanism inevitably results in temporal information loss. 

To validate our claim, we introduce a set of embarrassingly simple one-layer linear models named \emph{LTSF-Linear} for comparison. Experimental results on nine real-life datasets show that \emph{LTSF-Linear} surprisingly outperforms existing sophisticated Transformer-based LTSF models in all cases, and often by a large margin. Moreover, we conduct comprehensive empirical studies to explore the impacts of various design elements of LTSF models on their temporal relation extraction capability.
We hope this surprising finding opens up new research directions for the LTSF task. We also advocate revisiting the validity of Transformer-based solutions for other time series analysis tasks (e.g., anomaly detection) in the future. Code is available at: \url{https://github.com/cure-lab/LTSF-Linear}.

\end{abstract}


\input{section/Introduction}
\input{section/Related_Work}
\input{section/Method}
\input{section/DLinear}
\input{section/Experiment}
\input{section/Conclusion}
\pagebreak
% \widetext


\input{section/supp}
\input{main.bbl}
\bibliographystyle{ieee_fullname}


\end{document}
